---
title: "lod_no_food"
author: "guadalupe lulu bernal"
date: "2023-10-06"
output: word_document
---

```{r setup, include=FALSE}
# Set default options for all code chunks
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = 'center'
)
```

## R Markdown

## data analysis of LOD data food vs. no food with 2 strains for comparison, not a side by side comparison, food and no food trials were ran separate days with separate inoculum for each strain 

```{r load-libraries}
library(rstatix)
library(ggpubr)
library(broom)
library(skimr)
library(dplyr)
library(scales)
library(MASS)
library(ggplot2)
library(cowplot)
library(gtable)
require(MASS)
require(scales)
library(knitr)
library(tidyverse)
```

## add color palette

```{r color-palette}
cbPalette <- c("#999999", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
cbPalette_bluegreen <- c("#56B4E9", "#009E73")
```

## load file

```{r load-data}
lod_no_food <- read.csv("two_column_no_food.csv")
glimpse(lod_no_food)
```

## Create new column with log 10 transformed LOD

```{r transform-data}
### 
# df_lod_no_food <- lod_no_food  %>% 
#   group_by(device, food_matrix) %>%
#   glimpse()

df_lod_no_food <- lod_no_food  %>%
  mutate(log10_lod = log10(lod))


## use mutate here to change the labels so they are grammatically correct on the plot
mutate_df_lod_no_food <- df_lod_no_food %>%
  mutate(device = recode(device,
                           "decision_point" = "Decision Point",
                           "singlepath" = "SinglePath® Salmonella"))

## Change the df to the new df so it doesnt keep overwriting the df and undoing your work
mutate_df_lod_no_food <- mutate_df_lod_no_food %>%
  mutate(strain = recode(strain,
                           "S_Typhimurium" = "S. Typhimurium",
                           "S_Enteritidis" = "S. Enteritidis"))

mutate_df_lod_no_food <- mutate_df_lod_no_food %>%
  mutate(food_matrix = recode(food_matrix,
                           "absent " = "Absences of Food Matrix",
                           "present" = "Presence of Food Matrix"))

glimpse(mutate_df_lod_no_food)
```

## Exploratory plots

```{r exploratory-plots-1}
df_lod_no_food %>%
ggplot(aes(x = device, y = lod)) +
  geom_jitter(width = .2)
```


```{r exploratory-plots-2}
df_lod_no_food %>%
ggplot(aes(x = lod)) +
  geom_histogram(bins =8) +
  facet_grid(rows = vars(strain), scales = "free_y")
```

```{r exploratory-plots-3}
##
ggplot(df_lod_no_food, aes(y=lod, x=device, col = food_matrix)) +
  geom_crossbar(stat="summary", fun.data = mean_cl_normal, position = "dodge") +
  stat_summary(aes(group=device), fun = mean, geom="line") +
  geom_point(aes(group = device),size=2, position = position_jitterdodge()) +
       scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
     theme_bw() +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) 

## strains to lod for food matrix present 
ggplot(df_lod_no_food, aes(y=lod, x=strain, col = food_matrix)) +
  geom_crossbar(stat="summary", fun.data = mean_cl_normal, position = "dodge") +
  stat_summary(aes(group=device), fun = mean, geom="line") +
  geom_point(aes(group = device),size=2, position = position_jitterdodge()) +
       scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
     theme_bw() +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) 


## 

ggplot(df_lod_no_food, aes(y=lod, x=strain, col = food_matrix)) +
  geom_crossbar(stat="summary", fun.data = mean_cl_normal, position = "dodge") +
  stat_summary(aes(group=device), fun = mean, geom="line") +
  geom_point(aes(group = device),size=2, position = position_jitterdodge()) +
       scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
           labels = trans_format("log10", math_format(10^.x))) +
     theme_bw() +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) 

##facet plots
ggplot(df_lod_no_food, aes(y=lod, x=device, col = food_matrix)) +
  geom_crossbar(stat="summary", fun.data = mean_cl_normal, width = 0.2, col="black", fill = "gray") +
  geom_jitter(width=0.2, size = 2) +
  facet_grid(cols = vars(strain)) +
   theme_bw() +
       scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
     theme_bw() +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) 

## faceted by strain vs food matrix
ggplot(df_lod_no_food, aes(y=lod, x=food_matrix, col = device)) +
  geom_crossbar(stat="summary", fun.data = mean_cl_normal, width = 0.2, col="black", fill = "gray") +
  geom_jitter(width=0.2, size = 2) +
  facet_grid(cols = vars(strain)) +
   theme_bw() +
       scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
     theme_bw() +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) 

##
ggplot(df_lod_no_food, aes(y=lod, x=food_matrix, col = device)) +
  geom_crossbar(stat="summary", fun.data = mean_cl_normal, width = 0.2, col="black", fill = "gray") +
  facet_grid(cols = vars(strain)) +
   theme_bw() +
       scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
     theme_bw() +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) 

```
### log10_lod without log function

```{r log10-plots}
ggplot(df_lod_no_food, aes(y=log10_lod, x=device, col = food_matrix)) +
  geom_crossbar(stat="summary", fun.data = mean_cl_normal, position = "dodge") +
  stat_summary(aes(group=device), fun = mean, geom="line") +
  geom_point(aes(group = device),size=2, position = position_jitterdodge())
```


## Create data to only include one strain and changing labels for grammar/caps

```{r filter-strains}
## just S_Enteritidis
mutate_s.typ_lod <- mutate_df_lod_no_food %>%
    filter(strain == "S. Typhimurium")
glimpse(mutate_s.typ_lod)

## just S_Typhimurium
mutate_s.ent_lod <- mutate_df_lod_no_food %>%
  filter(strain == "S. Enteritidis")
glimpse(mutate_s.ent_lod)
```

####

```{r plot-typhimurium}
p <- ggplot(mutate_s.typ_lod, aes(device, log10_lod, col = food_matrix)) +
      scale_colour_manual(values=cbPalette_bluegreen) +
  geom_jitter() +
  facet_grid(cols = vars(food_matrix)) +
  xlab("Device") +
  ylab("LOD  CFU/mL")

strains_typ <- ggplot(mutate_s.typ_lod, aes(y=log10_lod, x=device, col=device)) +
        scale_colour_manual(values=cbPalette_bluegreen) +
  geom_jitter() +
  facet_grid(cols = vars(food_matrix)) +
#scale_y_log10(
#   breaks = scales::trans_breaks("log10", function(x) 10^x),
 #  labels = scales::trans_format("log10", scales::math_format(10^.x))
 #) +
    theme_bw() +
  theme(axis.text.x = element_text(size=10, angle=45, vjust=1, hjust=1)) +
  theme(axis.text.y = element_text(size=10)) +
  xlab("Device") +
  ylab("log10 of Limit of Detection  (CFU/mL)") +
   theme(axis.title = element_text(size = 10,
                                  face = "bold")) +
  theme(title = element_text(size = 12,
                                  face = "bold")) +
  theme(plot.title = element_text(lineheight=.8)) +
      guides(color = guide_legend(title = "Device")) +
      theme(axis.text.x=element_blank())
strains_typ
```

###

```{r plot-enteritidis}
p <- ggplot(mutate_s.ent_lod, aes(device, log10_lod, col = food_matrix)) +
      scale_colour_manual(values=cbPalette_bluegreen) +
  geom_jitter() +
  facet_grid(cols = vars(food_matrix)) +
  xlab("Device") +
  ylab("LOD  CFU/mL")

strains_ent <- ggplot(mutate_s.ent_lod, aes(y=log10_lod, x=device, col=device)) +
        scale_colour_manual(values=cbPalette_bluegreen) +
  geom_jitter() +
  facet_grid(cols = vars(food_matrix)) +
#scale_y_log10(
#   breaks = scales::trans_breaks("log10", function(x) 10^x),
 #  labels = scales::trans_format("log10", scales::math_format(10^.x))
 #) +
    theme_bw() +
  theme(axis.text.x = element_text(size=10, angle=45, vjust=1, hjust=1)) +
  theme(axis.text.y = element_text(size=10)) +
  xlab("Device") +
  ylab("log10 of Limit of Detection  (CFU/mL)") +
   theme(axis.title = element_text(size = 10,
                                  face = "bold")) +
  theme(title = element_text(size = 12,
                                  face = "bold")) +
  theme(plot.title = element_text(lineheight=.8)) +
      guides(color = guide_legend(title = "Device")) +
      theme(axis.text.x=element_blank())
strains_ent



```


##### paired t-test (2-tailed) to compare device sensitivity for all strains. No sig difference 

```{r summary-stats}
df_lod_no_food %>% 
  group_by(log10_lod) %>%
  summarise(mean_log10_lod = mean(log10_lod, na.rm = TRUE))


df_lod_no_food %>%  
  group_by(device) %>%
  summarise(mean_log10_lod = mean(log10_lod, na.rm = TRUE),
            sd_log10_lod = sd(log10_lod, na.rm = TRUE),
            n_log10_lod = n()) %>%
  mutate(SE_log10_lod = sd_log10_lod / sqrt(n_log10_lod),
         lower_95CI_t_dis = mean_log10_lod - qt(1 - (0.05 / 2), n_log10_lod - 1) * SE_log10_lod,
         upper_95CI_t_dis = mean_log10_lod + qt(1 - (0.05 / 2), n_log10_lod - 1) * SE_log10_lod)
```

### paired t-test (2-tailed) to compare device sensitivity for per stain.  No sig difference 

```{r lm-device}
lm_device_lod<- lm(lod ~ device, data = df_lod_no_food)
lm_device_lod


summary(lm_device_lod)
lm(formula = lod ~ device, data = df_lod_no_food)

confint(lm_device_lod)
```

### paired t-test (2-tailed) to compare device sensitivity for per stain no food.  No sig difference 

```{r summary-by-strain}
df_lod_no_food %>%  
  group_by(strain) %>%
  summarise(mean_log10_lod = mean(log10_lod, na.rm = TRUE),
            sd_log10_lod = sd(log10_lod, na.rm = TRUE),
            n_log10_lod = n()) %>%
  mutate(SE_log10_lod = sd_log10_lod / sqrt(n_log10_lod),
         lower_95CI_t_dis = mean_log10_lod - qt(1 - (0.05 / 2), n_log10_lod - 1) * SE_log10_lod,
         upper_95CI_t_dis = mean_log10_lod + qt(1 - (0.05 / 2), n_log10_lod - 1) * SE_log10_lod)

```

