---
title: "lod_no_food"
author: "guadalupe lulu bernal"
date: "2023-10-06"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

## data analysis of lod data food vs. no food with 2 strains for comparison, not a side by side comparison, food and no food trials were ran separate days with separate inoculmm for each strain 

```{r}
library(rstatix)
library(ggpubr)
library(broom)
library(skimr)
library(dplyr)
library(scales)
library(MASS)
library(ggplot2)
library(cowplot)
library(gtable)
require(MASS)
require(scales)
library(knitr)
library(tidyverse)
```

## add color palette

```{r}
cbPalette <- c("#999999", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
cbPalette_bluegreen <- c("#56B4E9", "#009E73")
```

## load file

```{r}
lod_no_food <- read.csv("two_column_no_food.csv")
glimpse(lod_no_food)




```

##create new column with log 10 transformed lod

```{r}
### 
# df_lod_no_food <- lod_no_food  %>% 
#   group_by(device, food_matrix) %>%
#   glimpse()

df_lod_no_food <- lod_no_food  %>%
  mutate(log10_lod = log10(lod))


## use mutate here to change the labels so they are grammatically correct on the plot
mutate_df_lod_no_food <- df_lod_no_food %>%
  mutate(device = recode(device,
                           "decision_point" = "Decision Point",
                           "singlepath" = "SinglePath® Salmonella"))

##change the df to the new df so it doesnt keep overwritting the df and undoing your work
mutate_df_lod_no_food <- mutate_df_lod_no_food %>%
  mutate(strain = recode(strain,
                           "S_Typhimurium" = "S. Typhimurium",
                           "S_Enteritidis" = "S. Enteritidis"))

mutate_df_lod_no_food <- mutate_df_lod_no_food %>%
  mutate(food_matrix = recode(food_matrix,
                           "absent " = "Absences of Food Matrix",
                           "present" = "Presence  of Food Matrix"))

glimpse(mutate_df_lod_no_food)
```

##exploratory plots

```{r}

df_lod_no_food %>%
ggplot(aes(x = device, y = lod)) +
  geom_jitter(width = .2)


##
df_lod_no_food %>%
ggplot(aes(x = lod)) +
  geom_histogram(bins =8) +
  facet_grid(rows = vars(strain), scales = "free_y")


## log scale histrogramm

df_lod_no_food %>%
ggplot(aes(x = log10_lod)) +
  geom_histogram(bins =8) +
  facet_grid(strain + device + food_matrix ~., scales = "free_y")
  

##


```

```{r}
##
ggplot(df_lod_no_food, aes(y=lod, x=device, col = food_matrix)) +
  geom_crossbar(stat="summary", fun.data = mean_cl_normal, position = "dodge") +
  stat_summary(aes(group=device), fun = mean, geom="line") +
  geom_point(aes(group = device),size=2, position = position_jitterdodge()) +
       scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
     theme_bw() +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) 

## strains to lod for food matrix present 
ggplot(df_lod_no_food, aes(y=lod, x=strain, col = food_matrix)) +
  geom_crossbar(stat="summary", fun.data = mean_cl_normal, position = "dodge") +
  stat_summary(aes(group=device), fun = mean, geom="line") +
  geom_point(aes(group = device),size=2, position = position_jitterdodge()) +
       scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
     theme_bw() +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) 


## 

ggplot(df_lod_no_food, aes(y=lod, x=strain, col = food_matrix)) +
  geom_crossbar(stat="summary", fun.data = mean_cl_normal, position = "dodge") +
  stat_summary(aes(group=device), fun = mean, geom="line") +
  geom_point(aes(group = device),size=2, position = position_jitterdodge()) +
       scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
           labels = trans_format("log10", math_format(10^.x))) +
     theme_bw() +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) 

##facet plots
ggplot(df_lod_no_food, aes(y=lod, x=device, col = food_matrix)) +
  geom_crossbar(stat="summary", fun.data = mean_cl_normal, width = 0.2, col="black", fill = "gray") +
  geom_jitter(width=0.2, size = 2) +
  facet_grid(cols = vars(strain)) +
   theme_bw() +
       scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
     theme_bw() +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) 

## faceted by strain vs food matrix
ggplot(df_lod_no_food, aes(y=lod, x=food_matrix, col = device)) +
  geom_crossbar(stat="summary", fun.data = mean_cl_normal, width = 0.2, col="black", fill = "gray") +
  geom_jitter(width=0.2, size = 2) +
  facet_grid(cols = vars(strain)) +
   theme_bw() +
       scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
     theme_bw() +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) 

##
ggplot(df_lod_no_food, aes(y=lod, x=food_matrix, col = device)) +
  geom_crossbar(stat="summary", fun.data = mean_cl_normal, width = 0.2, col="black", fill = "gray") +
  facet_grid(cols = vars(strain)) +
   theme_bw() +
       scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
     theme_bw() +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) 

```
### log10_lod without log function
```{r}
###
ggplot(df_lod_no_food, aes(y=log10_lod, x=device, col = food_matrix)) +
  geom_crossbar(stat="summary", fun.data = mean_cl_normal, position = "dodge") +
  stat_summary(aes(group=device), fun = mean, geom="line") +
  geom_point(aes(group = device),size=2, position = position_jitterdodge())

## strains to lod for food maxtrix present 
ggplot(df_lod_no_food, aes(y=log10_lod, x=strain, col = food_matrix)) +
  geom_crossbar(stat="summary", fun.data = mean_cl_normal, position = "dodge") +
  stat_summary(aes(group=device), fun = mean, geom="line") +
  geom_point(aes(group = device),size=2, position = position_jitterdodge())


## 

ggplot(df_lod_no_food, aes(y=log10_lod, x=strain, col = food_matrix)) +
  geom_crossbar(stat="summary", fun.data = mean_cl_normal, position = "dodge") +
  stat_summary(aes(group=device), fun = mean, geom="line") +
  geom_point(aes(group = device),size=2, position = position_jitterdodge())

##facet plots
ggplot(df_lod_no_food, aes(y=log10_lod, x=device, col = food_matrix)) +
  geom_crossbar(stat="summary", fun.data = mean_cl_normal, width = 0.2, col="black", fill = "gray") +
  geom_jitter(width=0.2, size = 2) +
  facet_grid(cols = vars(strain)) +
   theme_bw()

## faceted by strain vs food matrix
ggplot(df_lod_no_food, aes(y=log10_lod, x=food_matrix, col = device)) +
  geom_crossbar(stat="summary", fun.data = mean_cl_normal, width = 0.2, col="black", fill = "gray") +
  geom_jitter(width=0.2, size = 2) +
  facet_grid(cols = vars(strain)) +
   theme_bw()

##
ggplot(df_lod_no_food, aes(y=log10_lod, x=food_matrix, col = device)) +
  geom_crossbar(stat="summary", fun.data = mean_cl_normal, width = 0.2, col="black", fill = "gray") +
  facet_grid(cols = vars(strain)) +
   theme_bw()

```



## plotting food present vs. absent by strain

```{r}


p8 <- ggplot(mutate_df_lod_no_food, aes(y=lod, x=device, col = device)) +
  geom_crossbar(stat="summary", fun.data = mean_cl_normal, width = 0.2, col="black", fill = "gray") +
  geom_jitter(width=0.2, size = 2) +
  facet_grid(cols = vars(food_matrix)) +
   theme_bw() +
       scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
     theme_bw() +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1)) +
  ggtitle("Comparasion of LOD and Devices With and WithoutFood Matrix")


p9 <- ggplot(mutate_df_lod_no_food, aes(y=lod, x=food_matrix, col = device)) +
  geom_crossbar(stat="summary", fun.data = mean_cl_normal, width = 0.2, col="black", fill = "gray") +
  geom_jitter(width=0.2, size = 2) +
  facet_grid(cols = vars(strain)) +
   theme_bw() +
       scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
     theme_bw() +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x)))+
  theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1))  +
  ggtitle("Comparasion of LOD and Strains With and WithoutFood Matrix")

p5 <- plot_grid(p8,  p9, labels = c("a","b")) 
p5



```

## create data to only include one strain and changing labels for grammer/caps

```{r}


## just S_Enteritidis
mutate_s.typ_lod <- mutate_df_lod_no_food %>%
    filter(strain == "S. Typhimurium")
glimpse(mutate_s.typ_lod)

## just S_Typhimurium
mutate_s.ent_lod <- mutate_df_lod_no_food %>%
  filter(strain == "S. Enteritidis")
glimpse(mutate_s.ent_lod)


```

```{}
```

```{r}
p33 <- ggplot(mutate_s.ent_lod, aes(y=log10_lod, x=device)) +
  geom_crossbar(stat="summary", fun.data = mean_cl_normal, width = 0.2, col="black", fill = "gray") +
  geom_point() +
  facet_grid(cols = vars(food_matrix)) +
   theme_bw() +
theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1)) +
  xlab("Device") +
  ylab("LOD  CFU/mL")
p33 



p34 <- ggplot(mutate_s.ent_lod, aes(y=log10_lod, x=device)) +
  geom_crossbar(stat="summary", fun.data = mean_cl_normal, width = 0.2, col="black", fill = "gray") +
  geom_point() +
   theme_bw() +
theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1)) +
  xlab("Device") +
  ylab("LOD  CFU/mL")
p34

p36 <- plot_grid(p34, p33) 
p36

p35 <- ggplot(mutate_s.ent_lod, aes(y=log10_lod, x=food_matrix)) +
  geom_crossbar(stat="summary", fun.data = mean_cl_normal, width = 0.2, col="black", fill = "gray") +
  geom_point() +
   theme_bw() +
theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1)) +
  xlab("Food Matrix  (Ground Beef)") +
  ylab("LOD  Cfu/mL")
p35

p31 <- ggplot(mutate_s.typ_lod, aes(y=log10_lod, x=device)) +
  geom_crossbar(stat="summary", fun.data = mean_cl_normal, width = 0.2, col="black", fill = "gray") +
  geom_point() +
  facet_grid(cols = vars(food_matrix)) +
   theme_bw() +
theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1)) +
  xlab("Device") +
  ylab("LOD  CFU/mL")
p31



p32 <- ggplot(mutate_s.typ_lod, aes(y=log10_lod, x=device)) +
  geom_crossbar(stat="summary", fun.data = mean_cl_normal, width = 0.2, col="black", fill = "gray") +
  geom_point() +
   theme_bw()+
theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1)) +
  xlab("Device") +
  ylab("LOD  CFU/mL")
p32

p37 <- plot_grid(p31, p32) 
p37
 

p38 <- plot_grid(p37, p36,
                 nrow = 2,
  labels = "AUTO",
  label_size = 12,
  align = "v"
)
p38

 


```

```{r}
#compare_means(lod ~ device, data = mutate_df_lod_no_food, paired = TRUE)


cbPalette_bluegreen <- c("#56B4E9", "#009E73")

base <- ggpaired(mutate_df_lod_no_food, x = "device", y = "log10_lod",
         color = "device", line.color = "gray", line.size = 0.4) +
     scale_colour_manual(values=cbPalette_bluegreen)
 #      scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
#              labels = trans_format("log10", math_format(10^.x))) +
  theme(axis.text.x = element_text(size=10)) +
  theme(axis.text.y = element_text(size=10)) + 
  theme(axis.title.y = element_text(margin=margin(r=10))) +
  xlab("Food Matrix") +
  ylab("LOD  (CFU/mL)") +
 labs(title = stringr::str_wrap("Figure x. Comparasion of Salmonella Detection by Strain With and Without the Presence of Food", width=55)) +
  theme_bw() 

base + facet_grid(food_matrix ~ strain) +
      theme(axis.title.x = element_text(margin=margin(r=20))) +
 #     scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
#              labels = trans_format("log10", math_format(10^.x))) +
theme(axis.title = element_text(size = 10,
                                  face = "bold")) +
  theme(title = element_text(size = 12,
                                  face = "bold")) +
  theme(plot.title = element_text(lineheight=.8)) +
  theme(legend.position="none") 
theme_bw()


##### flip y and x axis


cbPalette_bluegreen <- c("#56B4E9", "#009E73")

base <- ggpaired(mutate_df_lod_no_food, x = "device", y = "log10_lod", color = "device") +
     scale_colour_manual(values=cbPalette_bluegreen)
 #      scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
#              labels = trans_format("log10", math_format(10^.x))) +
  theme(axis.text.x = element_text(size=10)) +
  theme(axis.text.y = element_text(size=10)) + 
  theme(axis.title.y = element_text(margin=margin(r=10))) +
  xlab("Food Matrix") +
  ylab("LOD  (CFU/mL)") +
 labs(title = stringr::str_wrap("Figure x. Comparasion of Salmonella Detection by Strain With and Without the Presence of Food", width=55)) +
  theme_bw()

base + facet_grid(strain ~ food_matrix) +
      theme(axis.title.x = element_text(margin=margin(r=20))) 
  #    scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
  #            labels = trans_format("log10", math_format(10^.x))) +
theme(axis.title = element_text(size = 10,
                                  face = "bold")) +
  theme(title = element_text(size = 12,
                                  face = "bold")) +
  theme(plot.title = element_text(lineheight=.8)) +
  theme(legend.position="none") 
```

#### device vs lod separated by strain

```{r}

p <- ggplot(mutate_s.typ_lod, aes(device, log10_lod, col = food_matrix)) +
      scale_colour_manual(values=cbPalette_bluegreen) +
geom_crossbar() +
  geom_point() +
  facet_grid(cols = vars(food_matrix)) +
  xlab("Device") +
  ylab("LOD  CFU/mL")

strains_typ <- ggplot(mutate_s.typ_lod, aes(y=log10_lod, x=device, col=device)) +
  geom_crossbar(stat="summary", fun.data = mean_cl_normal, width = 0.2, col="black", fill = "gray") +
        scale_colour_manual(values=cbPalette_bluegreen) +
  geom_point() +
  facet_grid(cols = vars(food_matrix)) +
   theme_bw() +
#scale_y_log10(
#   breaks = scales::trans_breaks("log10", function(x) 10^x),
 #  labels = scales::trans_format("log10", scales::math_format(10^.x))
 #) +
  xlab("Device") +
  ylab("LOD  CFU/mL")
strains_typ


 
```
####
```{r}

p <- ggplot(mutate_s.typ_lod, aes(device, log10_lod, col = food_matrix)) +
      scale_colour_manual(values=cbPalette_bluegreen) +
  geom_jitter() +
  facet_grid(cols = vars(food_matrix)) +
  xlab("Device") +
  ylab("LOD  CFU/mL")

strains_typ <- ggplot(mutate_s.typ_lod, aes(y=log10_lod, x=device, col=device)) +
        scale_colour_manual(values=cbPalette_bluegreen) +
  geom_jitter() +
  facet_grid(cols = vars(food_matrix)) +
#scale_y_log10(
#   breaks = scales::trans_breaks("log10", function(x) 10^x),
 #  labels = scales::trans_format("log10", scales::math_format(10^.x))
 #) +
    theme_bw() +
  theme(axis.text.x = element_text(size=10, angle=45, vjust=1, hjust=1)) +
  theme(axis.text.y = element_text(size=10)) +
  xlab("Device") +
  ylab("log10 of Limit of Detection  (CFU/mL)") +
   theme(axis.title = element_text(size = 10,
                                  face = "bold")) +
  theme(title = element_text(size = 12,
                                  face = "bold")) +
  theme(plot.title = element_text(lineheight=.8)) +
      guides(color = guide_legend(title = "Device")) +
      theme(axis.text.x=element_blank())
strains_typ
```

###

```{r}

p <- ggplot(mutate_s.ent_lod, aes(device, log10_lod, col = food_matrix)) +
      scale_colour_manual(values=cbPalette_bluegreen) +
  geom_jitter() +
  facet_grid(cols = vars(food_matrix)) +
  xlab("Device") +
  ylab("LOD  CFU/mL")

strains_ent <- ggplot(mutate_s.ent_lod, aes(y=log10_lod, x=device, col=device)) +
        scale_colour_manual(values=cbPalette_bluegreen) +
  geom_jitter() +
  facet_grid(cols = vars(food_matrix)) +
#scale_y_log10(
#   breaks = scales::trans_breaks("log10", function(x) 10^x),
 #  labels = scales::trans_format("log10", scales::math_format(10^.x))
 #) +
    theme_bw() +
  theme(axis.text.x = element_text(size=10, angle=45, vjust=1, hjust=1)) +
  theme(axis.text.y = element_text(size=10)) +
  xlab("Device") +
  ylab("log10 of Limit of Detection  (CFU/mL)") +
   theme(axis.title = element_text(size = 10,
                                  face = "bold")) +
  theme(title = element_text(size = 12,
                                  face = "bold")) +
  theme(plot.title = element_text(lineheight=.8)) +
      guides(color = guide_legend(title = "Device")) +
      theme(axis.text.x=element_blank())
strains_ent



```







###

```{r}

mutate_sum <- summarySE(mutate_df_lod_no_food, measurevar="log10_lod", groupvars=c("device","food_matrix"))
mutate_sum

# Error bars represent standard error of the mean
ggplot(mutate_sum, aes(x=device, y=log10_lod, fill=food_matrix)) + 
    geom_bar(position=position_dodge(), stat="identity") +
    geom_errorbar(aes(ymin=log10_lod-se, ymax=log10_lod+se),
                  width=.2,                    # Width of the error bars
                  position=position_dodge(.9))


# Use 95% confidence intervals instead of SEM
ggplot(mutate_sum, aes(x=device, y=log10_lod, fill=food_matrix)) + 
    geom_bar(position=position_dodge(), stat="identity") +
    geom_errorbar(aes(ymin=log10_LOD-ci, ymax=log10_lod+ci),
                  width=.2,                    # Width of the error bars
                  position=position_dodge(.9))
```



## absent vs. present of food

```{r}
stat.test <- mutate_df_lod_no_food %>%
  t_test(log10_lod ~ food_matrix) %>%
  adjust_pvalue() %>%
  add_significance()
stat.test

table_no_food_matrix <- kable(head(stat.test), format = "markdown", digits = 2)
table_no_food_matrix

# Box plots with p-values
bxp <-  ggplot(mutate_s.typ_lod, aes(y=log10_lod, x=device)) +
  geom_crossbar(stat="summary", fun.data = mean_cl_normal, width = 0.2, col="black", fill = "gray") +
  geom_point() +
  facet_grid(cols = vars(food_matrix)) 
stat.test <- stat.test %>% add_xy_position(x = "device")
  


# Customize p-value labels using glue expression 
# https://github.com/tidyverse/glue


```

#### downloading patterns to use oon ggplot <https://coolbutuseless.github.io/2020/04/01/introducing-ggpattern-pattern-fills-for-ggplot/>

```{r}
#install.packages("BiocManager")
#BiocManager::install("ArrayTools")
#install.packages("remotes")
```

#### 

```{r}
#remotes::install_github("coolbutuseless/ggpattern")

```

#### 

```{r}

```

### 

```{r}
library(ggsignif)
library(ggpattern)
df <- data.frame(level = c("Absent", "Present"), outcome = c(2.3, 1.9))
cbPalette_bluegreen <- c("#56B4E9", "#009E73")


mutate_df_lod_no_food %>% 
  ggplot(aes(x=device, y=log10_lod), pattern="stripe") +
  geom_col_pattern(aes(pattern_angle=food_matrix, color=food_matrix, fill=food_matrix),
                   pattern_fill="white", pattern_color="white", pattern_spacing=.05) +
    geom_signif(comparisons = list(c("absent", "present")), 
              map_signif_level=TRUE) +
  labs(x="Device", y="LOD") +
  facet_grid(~strain, switch = "x") +
  theme_classic()
###


###
gbar <- ggplot(mutate_df_lod_no_food,aes(x=device,y=log10_lod))+
  geom_bar(stat="identity",aes(fill=strain))
 
gbar+geom_signif(data=mutate_df_lod_no_food,stat="signif",position="identity",
                 comparisons=list(c("absent","present")),map_signif_level = TRUE,annotations="***")



```

#### plots  using log transformation to fix the y axis, but didn't work. 

```{r}

s.typ <-ggplot(mutate_s.typ_lod, aes(x = device, y = lod, pattern = food_matrix, fill = device)) +
geom_boxplot(aes(fill = device)) +
 #  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
#              labels = trans_format("log10", math_format(10^.x))) +
scale_fill_manual(name = "device", values = c("#56B4E9", "#009E73")) +
geom_boxplot_pattern(position = position_dodge(preserve = "single"), color = "black", pattern_fill = "white", pattern_angle = 45, pattern_density = 0.1, pattern_spacing = 0.025, pattern_key_scale_factor = 0.6) +
guides(pattern = guide_legend(override.aes = list(fill = "white")), fill = guide_legend(override.aes = list(pattern = "none"))) +
   labs(title = stringr::str_wrap("S. Typhimurium", width=55))

 #        scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
  #            labels = trans_format("log10", math_format(10^.x))) +
  theme(axis.text.x = element_text(size=10)) +
  theme(axis.text.y = element_text(size=10)) + 
  theme(axis.title.y = element_text(margin=margin(r=10))) +
  xlab("Device") +
  ylab("LOD  (CFU/mL)") +
  theme(axis.title = element_text(size = 10,
                                  face = "bold")) +
  theme(title = element_text(size = 12,
                                  face = "bold")) +
  theme(plot.title = element_text(lineheight=.8)) +
  theme_bw()
  
s.typ



s.ent <-ggplot(mutate_s.ent_lod, aes(x = device, y = lod, pattern = food_matrix, fill = device)) +
geom_boxplot(aes(fill = device)) +
 #    scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
  #            labels = trans_format("log10", math_format(10^.x))) +
scale_fill_manual(name = "food_matrix", values = c("#56B4E9", "#009E73")) +
    geom_jitter(width=0.2, size = 2) +
geom_boxplot_pattern(position = position_dodge(preserve = "single"), color = "black", pattern_fill = "white", pattern_angle = 45, pattern_density = 0.1, pattern_spacing = 0.025, pattern_key_scale_factor = 0.6) +
guides(pattern = guide_legend(override.aes = list(fill = "white")), fill = guide_legend(override.aes = list(pattern = "none"))) +
     labs(title = stringr::str_wrap("S. Enteritidis", width=55))

 #        scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
  #            labels = trans_format("log10", math_format(10^.x))) +
  theme(axis.text.x = element_text(size=10)) +
  theme(axis.text.y = element_text(size=10)) + 
  theme(axis.title.y = element_text(margin=margin(r=10))) +
  xlab("Device") +
  ylab("LOD  (CFU/mL)") +
  theme(axis.title = element_text(size = 10,
                                  face = "bold")) +
  theme(title = element_text(size = 12,
                                  face = "bold")) +
  theme(plot.title = element_text(lineheight=.8)) +
  theme_bw()
s.ent


pattern_boxplot <- plot_grid(s.typ, s.ent,
                 nrow = 2,
  labels = "AUTO",
  label_size = 12,
  align = "v"
)

pattern_boxplot 

```

####

```{r}

stat.test <- mutate_s.ent_lod %>%
  t_test(lod ~ food_matrix) %>%
  add_significance()
stat.test

plot_strains_pvalue <- ggboxplot(
  mutate_s.ent_lod, x = "device", y = "lod", 
  ylab = "LOD CFU/mL", xlab ="Device", add = "jitter"
  )

#bystrains

plot_strains_pvalue <- ggboxplot(
  mutate_s.ent_lod, x = "device", y = "lod", 
  ylab = "LOD CFU/mL", xlab ="Strain", add = "jitter"
  ) 

stat.test <- stat.test %>% add_xy_position(x = "device")
plot_strains_pvalue + 
  stat_pvalue_manual(stat.test, tip.length = 0) +
  labs(subtitle = get_test_label(stat.test, detailed = TRUE))




s.typ_plot <- ggplot(mutate_s.typ_lod, aes(y=log10_lod, x=device, color = device)) +
     scale_colour_manual(values=cbPalette_bluegreen) +
  geom_crossbar(stat="summary", fun.data = mean_cl_normal, width = 0.2, col="black", fill = "gray") +
  geom_jitter(width=0.2, size = 1.5) +
  facet_grid(cols = vars(food_matrix))  +
    theme_classic() +
  theme(axis.text.x = element_text(size=10, angle=45, vjust=1, hjust=1)) +
  theme(axis.text.y = element_text(size=10)) +
  xlab("Device") +
  ylab("log10 of Limit of Detection  (CFU/mL)") +
   theme(axis.title = element_text(size = 10,
                                  face = "bold")) +
  theme(title = element_text(size = 12,
                                  face = "bold")) +
  theme(plot.title = element_text(lineheight=.8)) +
      guides(color = guide_legend(title = "Device")) +
      theme(axis.text.x=element_blank())
s.typ_plot



ggplot(mutate_s.typ_lod, aes(y = log10_lod, x = device)) +
  geom_jitter(width=0.1, size = 2, alpha = 0.5, 
              fill = "lightgreen", shape = 21) +
  geom_crossbar(stat = "summary", fun.data = mean_cl_normal, 
                width = 0.2, col="black", fill = "gray", 
                alpha = 0.5) +
  theme_bw() +
  xlab("Device") +
  ylab("log10 of Limit of Detection (CFU/mL)")


s.ent_plot <- ggplot(mutate_s.ent_lod, aes(y=log10_lod, x=device, color = device)) +
     scale_colour_manual(values=cbPalette_bluegreen) +
  geom_crossbar(stat="summary", fun.data = mean_cl_normal, width = 0.2, col="black", fill = "gray") +
  geom_jitter(width=0.2, size = 1.5) +
  facet_grid(cols = vars(food_matrix))  +
    theme_classic() +
  theme(axis.text.x = element_text(size=10, angle=45, vjust=1, hjust=1)) +
  theme(axis.text.y = element_text(size=10)) +
  xlab("Device") +
  ylab("log10 of Limit of Detection  (CFU/mL)") +
   theme(axis.title = element_text(size = 10,
                                  face = "bold")) +
  theme(title = element_text(size = 12,
                                  face = "bold")) +
  theme(plot.title = element_text(lineheight=.8)) +
      guides(color = guide_legend(title = "Device")) +
      theme(axis.text.x=element_blank())
s.ent_plot
```
###
```{r}
s.ent_plot <- ggplot(mutate_s.ent_lod, aes(y=log10_lod, x=device, color = device)) +
     scale_colour_manual(values=cbPalette_bluegreen) +
  geom_jitter(width=0.2, size = 1.5) +
  facet_grid(cols = vars(food_matrix))  +
    theme_classic() +
  theme(axis.text.x = element_text(size=10, angle=45, vjust=1, hjust=1)) +
  theme(axis.text.y = element_text(size=10)) +
  xlab("Device") +
  ylab("log10 of Limit of Detection  (CFU/mL)") +
   theme(axis.title = element_text(size = 10,
                                  face = "bold")) +
  theme(title = element_text(size = 12,
                                  face = "bold")) +
  theme(plot.title = element_text(lineheight=.8)) +
      guides(color = guide_legend(title = "Device")) +
      theme(axis.text.x=element_blank())
s.ent_plot
```



#### Table

```{r}

# BUSINESS SCIENCE R TIPS ----
# R-TIP 034 | mmtable2: ggplot2 syntax for tables ----
#
# 👉 For Weekly R-Tips, Sign Up Here: https://learn.business-science.io/r-tips-newsletter

# LIBRARIES ----

#remotes::install_github("ianmoran11/mmtable2")

#library(mmtable2)
#library(gt)
#library(tidyverse)

#mpg


# 1.0 Data Transformation -----

#data_wrangled <- mutate_df_lod_no_food %>%
 #   group_by(manufacturer, cyl) %>%
 #   summarise(across(.cols = c(cty, hwy), .fns = mean)) %>%
  #  ungroup() %>%
 #   pivot_longer(
  #      cols = c(cty, hwy),
 #       names_to = "fuel_economy_type",
 #       values_to = "fuel_economy"
 #   )

#data_wrangled


# 2.0 Table Main ----
#main_table <- data_wrangled %>%
#    mutate(fuel_economy = round(fuel_economy, 1)) %>%

#    mmtable(cells = fuel_economy, table_name = "Fuel Economy") +

    # Specify Headers
#    header_top(manufacturer) +
 #   header_left(cyl) +
 #   header_left_top(fuel_economy_type) +
#

    # Specify formatting
#    header_format(manufacturer, list(cell_text(transform = "capitalize"))) +
#    header_format(fuel_economy_type, list(cell_text(transform = "uppercase"))) +
#    table_format(
#        locations = list(
 #           cells_body(rows = c(2, 6))
#        ),
 #       style     = list(
 #           cell_borders(sides = "top", color = "grey")
 #       )
  #  )

# 3.0 Modify with GT Table ----
#main_table %>%
#    gt::tab_header(
##        title = "Fuel Economy by Car Manufacturer",
 #       subtitle = "Audi, VW, and Honda are leaders in Fuel Economy"
#    )

# LEARNING MORE ----

# FREE MASTERCLASS
# - 10 SECRETS TO BECOMING A DATA SCIENTIST
#   https://learn.business-science.io/free-rtrack-masterclass



```

### separate charts for strains, first new data frames

```{r}


subset_s.t_a <- subset(mutate_df_lod_no_food, strain == 'S. Typhimurium' & food_matrix %in% c('Absent'))
subset_s.t_p <- subset(mutate_df_lod_no_food, strain == 'S. Typhimurium' & food_matrix %in% c('Present'))


subset_s.e_a <- subset(mutate_df_lod_no_food, strain == 'S. Enteritidis' & food_matrix %in% c('Absent'))

subset_s.e_p <- subset(mutate_df_lod_no_food, strain == 'S. Enteritidis' & food_matrix %in% c('Present'))

```

### S. Typhimurium, devices vs. lod by absent/present

#### 

```{r}

plot_sta <- ggplot(subset_s.t_a, aes(y=log10_lod, x=device)) +
  geom_crossbar(stat="summary", fun.data = mean_cl_normal, width = 0.2, col="black", fill = "gray") +
  geom_point() +
  facet_grid(cols = vars(food_matrix)) +
   theme_bw() +
theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1)) +
  xlab("Device") +
  ylab("log10 (LOD)  CFU/mL") 
plot_sta
```

## sample data to see what the table looks like

```{r}
data(airquality)

# creates a table of the first six (6) rows of "airquality"
kable(head(airquality), format = "markdown", digits = 2)

```

## stats analysis <https://www.r-bloggers.com/2017/06/add-p-values-and-significance-levels-to-ggplots/>

```{r}
stat.test <- mutate_df_lod_no_food %>%
  group_by(strain, food_matrix) %>%
  t_test(log10_lod ~ device) %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_significance()
stat.test

###
stat.test <- mutate_df_lod_no_food %>%
  group_by(strain, food_matrix) %>%
  t_test(lod ~ device) %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_significance()
stat.test

##


#stat.test <- stat.test %>% add_xy_position(x = "device")
#ggboxplot(
#  mutate_df_lod_no_food, x = "device", y = "log10_lod", fill = "#00AFBB",
#  facet = c("strain", "food_matrix")
#  ) +
#  stat_pvalue_manual(
#    stat.test, bracket.nudge.y = -2, hide.ns = TRUE,
#    label = "{p.adj}{p.adj.signif}"
#    ) +
#  background_grid(major = 'y', minor = "x")



###
compare_means(lod ~ device, data = mutate_df_lod_no_food, paired = TRUE)

ggpaired(mutate_s.typ_lod, x = "food_matrix", y = "log10_lod",
         color = "device", line.color = "gray", line.size = 0.4,
         palette = "jco")+
  stat_compare_means(paired = TRUE)


#compare_means(lod ~ food_matrix, data = mutuate_s.ent_lod, paired = TRUE)

##
data(stat.test)

# creates a table of the first six (6) rows of "airquality"
kable(head(stat.test), format = "markdown", digits = 2)

```

## plots per strain

```{r}

## just S_Typhimurium
 p1 <- ggplot(mutate_s.typ_lod, aes(y=log10_lod, x=device)) +
    geom_crossbar(stat="summary", fun.data = mean_cl_normal, width = 0.2, col="black", fill = "gray") +
  geom_jitter(width=0.2, size = 2) +
  facet_grid(cols = vars(food_matrix)) +
   theme_bw() +
theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1)) +
  xlab("Device") +
  ylab("LOD  CFU/mL") +
 ggtitle("S. Typhimurium")

p1
 
## just S_Enteritidis

 p2 <- ggplot(mutate_s.ent_lod, aes(y=log10_lod, x=device)) +
    geom_crossbar(stat="summary", fun.data = mean_cl_normal, width = 0.2, col="black", fill = "gray") +
  geom_jitter(width=0.2, size = 2) +
  facet_grid(cols = vars(food_matrix)) +
   theme_bw() +
  xlab("Device") +
  ylab("LOD  CFU/mL") +
 ggtitle("S. Ent")

 


 
p3 <- plot_grid(p1,  p2, labels = c("a","b"))
p3

### adding log scale for better y axis 

require(scales)


```
#####

```{r}



## faceted by strain vs food matrix
strain_food_1 <- ggplot(mutate_df_lod_no_food, aes(y=log10_lod, x=food_matrix, col = device)) +
  geom_crossbar(stat="summary", fun.data = mean_cl_normal, width = 0.2, col="black", fill = "gray") +
  geom_jitter(width=0.2, size = 2) +
  facet_grid(cols = vars(strain)) +
   theme_bw() +
  theme(axis.text.x = element_text(size=10, angle=45, vjust=1, hjust=1)) +
  theme(axis.text.y = element_text(size=10)) +
  xlab("Food Matrix") +
  ylab("LOD  CFU/mL") +
   theme(axis.title = element_text(size = 10,
                                  face = "bold")) +
  theme(title = element_text(size = 12,
                                  face = "bold")) +
  theme(plot.title = element_text(lineheight=.8)) +
 labs(title = stringr::str_wrap("Figure 2. Comparasion of Salmonella Detection by Strain With and Without The Presence of Food Matrix", width=55)) +
  theme(legend.position="none")
ggsave("Figure X. Comparasion of Salmonella Detection by Strain With and Without The Presence of Food Matrix.png")

strain_food_1




```

#### 

```{r}
### trying to separate the previous plot into 4 graphs instead, 2 for device, 2 for food matrix. these are confusing to look at through 

## separated by device vs food matrix
ggplot(data = mutate_df_lod_no_food, aes(food_matrix, log10_lod, color = strain)) +
  geom_boxplot() + 
    theme_bw() +
  theme(axis.text.x = element_text(size=10, angle=45, vjust=1, hjust=1)) +
  theme(axis.text.y = element_text(size=10)) +
  xlab("Food Matrix") +
  ylab("LOD  CFU/mL") +
   theme(axis.title = element_text(size = 10,
                                  face = "bold")) +
  theme(title = element_text(size = 12,
                                  face = "bold")) +
  theme(plot.title = element_text(lineheight=.8)) +
  theme_bw() +
  labs(title = "Figure x. Comparasion of Salmonella Detection by Strain With and Without The Presence of Food Matrix",
       subtitle = "p = .XXX",
       y = "food matrix", x = "log10(LOD) (cfu/mL)") + 
  facet_wrap(~ food_matrix + strain)


```

```{r}
##distribution of lod by device

df_lod_no_food %>% 
  group_by(log10_lod) %>%
  summarise(mean_log10_lod = mean(log10_lod, na.rm = TRUE))


df_lod_no_food %>%  
  group_by(device) %>%
  summarise(mean_log10_lod = mean(log10_lod, na.rm = TRUE),
            sd_log10_lod = sd(log10_lod, na.rm = TRUE),
            n_log10_lod = n()) %>%
  mutate(SE_log10_lod = sd_log10_lod / sqrt(n_log10_lod),
         lower_95CI_t_dis = mean_log10_lod - qt(1 - (0.05 / 2), n_log10_lod - 1) * SE_log10_lod,
         upper_95CI_t_dis = mean_log10_lod + qt(1 - (0.05 / 2), n_log10_lod - 1) * SE_log10_lod)



##

```

### week 7 from biostats making a linear model for prettier plots?

```{r}

lm_device_lod<- lm(lod ~ device, data = df_lod_no_food)
lm_device_lod


summary(lm_device_lod)
lm(formula = lod ~ device, data = df_lod_no_food)

confint(lm_device_lod)



```

### 

#### log scales instead

```{r}

log_plot <- ggplot(df_lod_no_food, aes(x = device, y = lod)) + 
  geom_boxplot()




require(scales)
log_plot + scale_y_continuous(trans = log2_trans(),
    breaks = trans_breaks("log2", function(x) 2^x),
    labels = trans_format("log2", math_format(2^.x)))
```

### 

```{r}

log_plot <- ggplot(df_lod_no_food, aes(x = device, y = lod)) + 
  geom_boxplot()




require(scales)
log_plot + scale_y_continuous(trans = log2_trans(),
    breaks = trans_breaks("log2", function(x) 2^x),
    labels = trans_format("log2", math_format(2^.x)))
```

### 

```{r}

#plot_strains_pvalue <- ggboxplot(
#  f_lod_food, x = "device", y = "lod", 
 # ylab = "LOD CFU/mL", xlab ="Device", add = "jitter"
 # )

#stat.test <- stat.test %>% add_xy_position(x = "device")
#plot_strains_pvalue + 
 #stat_pvalue_manual(stat.test, tip.length = 0) +
#labs(subtitle = get_test_label(stat.test, detailed = TRUE))


```

## stats analysis

```{r}
t.test(lod ~ food_matrix, data = df_lod_no_food)

## lod vs. strain
stat.test <- df_lod_no_food  %>% 
  t_test(lod ~ strain, paired = TRUE) %>%
  add_significance()
stat.test




```

## 

```{r}
plot_pvalue <- ggboxplot(
  mutuate_f_lod_no_food, x = "food_matrix", y = "lod", 
  ylab = "LOD CFU/mL", xlab ="Strain", add = "jitter"
  ) 


stat.test <- stat.test %>% add_xy_position(x = "food_matrix")
plot_pvalue + 
 scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x)))




```

## 

```{r}

#dishonest_jitter <- ggplot(df_lod_no_food, aes(x = food_matrix, y = lod))  + 
 # geom_jitter()+
 # labs(title = "dishonest jitter")

#honest_jitter <- ggplot(df_lod_no_food, aes(x = cyandensity, y = lod, color =  strain))  + 
#  geom_jitter(height = 0, width = .2, show.legend = FALSE)+
 # labs(title = "honest jitter")

#plot_grid(dishonest_jitter, honest_jitter, labels = c("a","b"))



```
